```{title: "Close rushing wins"}
author: "Eli Cohen"
date: '2024-10-23'
format: 
  html:
    code-fold: true
```


```{r}
library(tidyverse)
```

```{r}
ggplot() +
 geom_line(data = BigTen, aes(x = Season, y = Win_pct, color = "Big Ten")) +
  geom_point(data = BigTen, aes(x = Season, y = Win_pct, color = "Big Ten")) +
  geom_line(data = SEC, aes(x = Season, y = Win_pct, color = "SEC")) +
  geom_point(data = SEC, aes(x = Season, y = Win_pct, color = "SEC")) +
  geom_line(data = Big12, aes(x = Season, y = Win_pct, color = "Big 12")) +
  geom_point(data = Big12, aes(x = Season, y = Win_pct, color = "Big 12")) +
  scale_x_continuous(breaks = seq(2015, 2024, by = 3))+

  scale_color_manual(values = c("Big Ten" = "black", 
                                 "SEC" = "maroon", 
                                 "Big 12" = "navy")) +
  labs(color = "Conference",
       title = "No run game? No problem.",
       caption = "The likelihood of winning a game without a strong run game has increased in the last decade. ",
       y = "Winning percentage",
       subtitle = "Team performance when rushing for less than 100 yards.") +
  theme_minimal() 

```

Once again, it looks like the "run the damn ball" guy is wrong again aside from an outlier from the Big 12 in 2022. It reinforces the idea that quarterback play is king, and shows why most of the money goes to whoever's under center. Even in the Big Ten — which was dominated by a Jim Harbaugh-led Michigan team that ran it as well as anyone in the country — a bad game running the ball has became increasingly-less important in recent years.

```{r}

logs <- footballlogs1124_3_
```

```{r}
 winlosslogs <- logs |> 
  mutate(
    wins = case_when(
      grepl("W", Outcome) ~ 1, 
      grepl("L", Outcome) ~ 0)
  )
```

```{r}
small_rushing_wins <- winlosslogs |>
filter(RushingYds<=100 & wins == 1)

small_rushing <- winlosslogs |>
  filter(RushingYds<=100)
```

```{r}
data <- (small_rushing) |> group_by(Season,Conference) |> summarize(Games = n(), Wins=sum(wins), Win_pct=Wins/Games) |>
filter(Season>="2015")

BigTen <- data |> filter(Conference == "Big Ten Conference")
SEC <- data |> filter(Conference =="Southeastern Conference")
CUSA <- data |> filter(Conference =="Conference USA")
Big12 <- data |> filter(Conference =="Big 12 Conference")
Ind <- data |> filter(Conference =="Independent")
MAC <- data |> filter(Conference =="Mid-American Conference")
PAC12 <- data |> filter(Conference =="Pac-12 Conference"| Conference == "Pacific-12 Conference")
BigEast <- data |> filter(Conference =="Big East Conference")
SunBelt <- data |> filter(Conference =="Sun Belt Conference")
MWC <- data |> filter(Conference =="Mountain West Conference")
WAC <- data |> filter(Conference =="Western Athletic Conference")
ACC <- data |> filter(Conference == "Atlantic Coast Conference")
```
